Meta-analytic approaches and effect sizes to account for 'nuisance heterogeneity' in comparative physiology

### 2021-09-17 MS/ms.Rmd
Fixing figure. Some small changes in levels in data.

```{r, figsurv, fig.wid = 16.833333, fig.height=5,fig.cap="**Summary of meta-analytic practices in comparative physiology.** (A) Different types of effect measures commonly used in meta-analyses and their relative frequency of use. (B) Frequency of studies where ‘nuisance heterogenity’ had the potential to impact effect sizes (x-axis), and the frequencies of studies explicitly dealing with nuisance heterogeneity (bar colour). (C) Frequencies of analytical approaches used to analyse effect measures. Note that only ‘Weighted Regression’, ‘Random Effects Models’, ‘Multilevel Metaregression Models’ and ‘Multilevel Meta-analysis Models’ are weighted models. The results of this survey were drawn from 62 studies in comparative physiology."}
p1 = dat %>% dplyr::group_by(effect_size_type) %>% janitor::tabyl(effect_size_type)  %>% 
  dplyr::summarise(SMD = dplyr::if_else(grepl("SMD", effect_size_type), n, 0),
                  lnRR = dplyr::if_else(grepl("Response Ratio", effect_size_type), n, 0),
                    Zr = dplyr::if_else(grepl("correlation", effect_size_type), n, 0),
                 Other = dplyr::if_else(grepl("other|Q10", effect_size_type), n, 0),
                 Mean = dplyr::if_else(grepl("mean/proportion/raw", effect_size_type), n, 0),
                 lnCVR = dplyr::if_else(grepl("lnCVR", effect_size_type), n, 0),
                 Slope = dplyr::if_else(grepl("slope", effect_size_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = n / sum(n)) %>% arrange(effect) %>% 
  mutate(effect = fct_relevel(effect, c("Other", "Zr", "lnRR", "SMD", "lnCVR", "Slope", "Means"))) %>% 
  ggplot() + tm + 
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = "black", colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,0.35)) +
  labs(x = "Effect Size",
       y = "Frequency") +
  coord_flip()

@@

p3 <- summary %>% 
  mutate(effect = fct_relevel(effect, c("Other","Vote Counting", "Linear Regression", "Weighted Regression", "Random-Effects Model", "Multilevel Meta-analysis Model", "Multilevel Metaregression Model"))) %>% 
  ggplot() + tm +
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = "black", colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),limits = c(0,0.4)) +
  labs(x = "Analytical Approach",
       y = "Frequency") +
  coord_flip()

@@

Heterogeneity described using $I^2$ measures can be useful for understanding the relative contributions of different factors to effect size variation; however, prediction intervals might be more appropriate in many cases [@Nakagawa2021c; @Borenstein2019]. Unlike $I^2$, prediction intervals (PI's) are probably more meaningful in the context of meta-analysis because they explicitly incorporate measures of dispersion around a mean effect size. They provide information about the likely effect size one can expect if we were to randomly sample a new population. For example, assume that we were interested in the overall impact of salinity stress in freshwater fish. To tackle this question, we collected experimental studies measuring fish swim performance under high salinity treatments (\~15--20ppm Na Cl) compared to freshwater (\~0ppm Na Cl) controls, using *SMD* to quantify the effect salinity had on swim performance. We conduct a meta-analysis and estimate an overall *SMD* of 0.50 with a 95% prediction interval of 0.05 to 0.85. The PI indicates that effects can vary widely from as low as an *SMD* of 0.05 to as high as an *SMD* 0.85, depending on the population. In addition, if we were to repeat a similar study, we would expect a new effect to fall within the range of 0.05 and 0.85, 95% of the time [@Borenstein2019; @Nakagawa2021c]. This is a very intuitive interpretation of how variable effects are, yet PIs are often not reported in meta-analyses [@Nakagawa2021c], and we encourage more meta-analysts to report these (for examples of PIs, see Supplementary Materials).

>>>>

```{r, figsurv, fig.wid = 16.833333, fig.height=5,fig.cap="**Summary of meta-analytic practices in comparative physiology.** (A) Different types of effect measures commonly used in meta-analyses and their relative frequency of use. (B) Frequency of studies where ‘nuisance heterogenity’ had the potential to impact effect sizes (x-axis), and the frequencies of studies explicitly dealing with nuisance heterogeneity (bar colour). (C) Frequencies of analytical approaches used to analyse effect measures. Note that only ‘Weighted Regression’, ‘Random Effects Models’, ‘Multilevel Metaregression Models’ and ‘Multilevel Meta-analysis Models’ are weighted models. The results of this survey were drawn from 62 studies in comparative physiology."}
p1 = dat %>% dplyr::group_by(effect_size_type) %>% janitor::tabyl(effect_size_type)  %>% 
  dplyr::summarise(SMD = dplyr::if_else(grepl("SMD", effect_size_type), n, 0),
                  lnRR = dplyr::if_else(grepl("Response Ratio", effect_size_type), n, 0),
                    Zr = dplyr::if_else(grepl("correlation", effect_size_type), n, 0),
                 Other = dplyr::if_else(grepl("other|Q10", effect_size_type), n, 0),
                 Mean = dplyr::if_else(grepl("mean/proportion/raw", effect_size_type), n, 0),
                 lnCVR = dplyr::if_else(grepl("lnCVR", effect_size_type), n, 0),
                 Slope = dplyr::if_else(grepl("slope", effect_size_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = n / sum(n)) %>% arrange(effect) %>% 
  mutate(effect = fct_relevel(effect, c("Other", "Zr", "lnRR", "SMD", "lnCVR", "Slope", "Mean"))) %>% 
  ggplot() + tm + 
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = "black", colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,0.35)) +
  labs(x = "Effect Size",
       y = "Frequency") +
  coord_flip()

@@

p3 <- summary %>% 
  mutate(effect = fct_relevel(effect, c("Other","Vote Counting", "Linear Regression", "Weighted Regression", "Random Effects Model", "Multilevel Meta-analysis Model", "Multilevel Metaregression Model"))) %>% 
  ggplot() + tm +
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = "black", colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),limits = c(0,0.4)) +
  labs(x = "Analytical Approach",
       y = "Frequency") +
  coord_flip()

@@

Heterogeneity described using $I^2$ measures can be useful for understanding the relative contributions of different factors to effect size variation; however, prediction intervals might be more appropriate in many cases [@Nakagawa2021c; @Borenstein2019]. Unlike $I^2$, prediction intervals (PI's) are probably more meaningful in the context of meta-analysis because they explicitly incorporate measures of dispersion around a mean effect size. They provide information about the likely effect size one can expect if we were to randomly sample a new population. For example, assume that we were interested in the overall impact of salinity stress in freshwater fish. To tackle this question, we collected experimental studies measuring fish swimming performance under high salinity treatments (\~15--20ppm Na Cl) compared to freshwater (\~0ppm Na Cl) controls, using *SMD* to quantify the effect salinity had on swim performance. We conduct a meta-analysis and estimate an overall *SMD* of 0.50 with a 95% prediction interval of 0.05 to 0.85. The PI indicates that effects can vary widely from as low as an *SMD* of 0.05 to as high as an *SMD* 0.85, depending on the population. In addition, if we were to repeat a similar study, we would expect a new effect to fall within the range of 0.05 and 0.85, 95% of the time [@Borenstein2019; @Nakagawa2021c]. This is a very intuitive interpretation of how variable effects are, yet PIs are often not reported in meta-analyses [@Nakagawa2021c], and we encourage more meta-analysts to report these (for examples of PIs, see Supplementary Materials).
